---
title: "Practica"
author: "Sebastian Ortuno"
date: "2025-06-21"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this assignment, I will build a DNN model based on the IRIS dataset based on the pdf I found in BB. I made sure first to install all the libraries needed for this project before I start.

I divided this project in the following steps:

1. Load the libraries.
2. 

# **1. Load the libraries**


```{r}
library(keras) #Load Keras Library
library(tensorflow) #Load Tensorflow
# Set seed for reproducibility
tensorflow::set_random_seed(1)
```

# **2. Prepare the dataset**  

```{r}
data <- iris[sample(nrow(iris)), ]  # shuffle rows and randomize order of samples.
x <- data[, 1:4] #This separates the features
y <- data$Species #Separates  species column.

# Normalize features to [0,1]
x <- as.matrix(apply(x, 2, function(col) (col - min(col)) / (max(col) - min(col))))


y <- to_categorical(as.integer(as.factor(y)) - 1, num_classes = 3) #  Convert lables to integers and then apple One-hot encoding.
```

# **3. Build the model.**

```{r}

model <- keras_model_sequential() %>% #Define a sequential DNN model.
  layer_dense(units = 10, activation = "relu", input_shape = ncol(x)) %>% #Input plus first hidden layer
  layer_dense(units = 10, activation = "relu") %>% #Second hidden layer
  layer_dense(units = 3, activation = "softmax") #output layer(3 classes)
```


# **4. Compile the model.**

```{r}
model %>% compile( #Compile function
  loss = "categorical_crossentropy",
  optimizer = "adam", #efficient adaptive optimizer. 
  metrics = "accuracy" #This track the accuracy during training.
)

```

# **5. Train the model**

```{r}
fit <- model %>% fit( #train the model
  x = x, #Features
  y = y, #one-hot encoded labels. 
  shuffle = TRUE, #Shuffle data before each epoch
  batch_size = 5, #batch size
  validation_split = 0.3, #30% for validation
  epochs = 200 #Number of training epochs
)

```


# **6.Evaluate performance**

```{r}
score <- model %>% evaluate(x, y) #Evaluate the model on the dataset
cat("Test accuracy:", score["accuracy"], "\n") #Print the test accuracy
```

# **7. Graphics**

```{r}
# Plot training history
plot(fit)

```